---
title: "HowWarsEnd"
output: html_document
---


```{r}
library(dplyr)
```


```{r warning=FALSE}
#install.packages("tidyverse")
termination_data = readr::read_csv(here::here("data", "raw", "UCDPConflictTerminationDataset_v4_2024_Conflict.csv"))

conflict_data = readr::read_csv(here::here("data", "raw", "UcdpPrioConflict_v25_1.csv"))
```

##### Variables of interest: 

**type of conflict2**
The same conflict episode, or dyadic conflict episode, may include both years
where neither side receive secondary support and years when they do. Type
2 thus combine the categories of internal armed conflict and internationalized
armed conflict into the categories described below
1. Extrasystemic armed conflict.
2. Interstate armed conflict.
3. Intrastate armed conflict.


**c outcome**
The coding of outcomes are based on the final year of activity and first year of
non-activity.
1= Peace agreement
2= Ceasefire
3= Victory for Side A /Government Side
4= Victory for Side B /Non-State Side
5= Low activity (some, but less than 25 battle-deaths in subsequent year(s))
6= Actor ceases to exist



```{r}
termination_data <- termination_data %>%
  mutate(c_outcome_simple = case_when(
    is.na(c_outcome)           ~ "ongoing",
    c_outcome %in% c(1, 2)     ~ "diplomatic",
    c_outcome %in% 3:6         ~ "military",
    TRUE                       ~ "other"  # Catch-all for unexpected values
  ))
```

```{r}
# create decade variable and termination variable
termination_data <- termination_data %>%
  mutate(decade = (year %/% 10) * 10) 


termination_data <- termination_data %>%
  # Sort by ID and Year to ensure 'lead' picks the correct next row
  arrange(conflict_id, year) %>%
  
  # Group by ID so we only compare years within the same conflict
  group_by(conflict_id) %>%
  
  # Create the duration (gap) variable
  mutate(duration = ifelse(
    c_epterm == 1, 
    lead(year) - year, 
    NA
  )) %>%
  ungroup()
```


```{r}
termination_data <- termination_data %>%
  # Ensure data is ordered correctly by conflict and time
  arrange(conflict_id, year) %>%
  group_by(conflict_id) %>%
  mutate(
    # 1. Create conflict_resumes
    # Check if any subsequent year exists for this conflict_id
    conflict_resumes = ifelse(
      c_epterm == 1, 
      as.integer(year < max(year)), 
      NA
    ),
    # 2. Create peace_duration
    # Calculate years until the next recorded observation for this conflict
    peace_duration = ifelse(
      c_epterm == 1 & conflict_resumes == 1,
      lead(year) - year,
      NA
    )
  ) %>%
  ungroup()

termination_data <- termination_data %>%   
  mutate(
    # 1. Create conflict_resumes
    # Check if any subsequent year exists for this conflict_id
    peace_never_fails = ifelse(
      c_epterm == 1 & is.na(peace_duration), 
      as.integer(max(year)-year), 
      NA
    )
  )

termination_data <- termination_data %>%   
  mutate(
    # 1. Create conflict_resumes
    # Check if any subsequent year exists for this conflict_id
    peace_fails_in_10 = ifelse(
      c_epterm == 1 & (peace_duration>10 | is.na(peace_duration)), 
      0, 
      1
    )
  )

```

```{r}
library(ggplot2)

ggplot(termination_data %>% filter(type_of_conflict %in% c(2, 4)), 
       aes(x = year, fill = c_outcome_simple)) +
  geom_bar(position = "stack") +
  scale_x_continuous(
    breaks = seq(from = 1900, to = 2020, by = 10) # Adjust start/end to your data
  ) +
  labs(
    title = "War Terminations Each Year by Outcome",
    x = "Year",
    y = "Wars Terminated",
    fill = "Outcome Type"
  ) +
  theme_minimal()


```



```{r}
# Pre-calculate the counts per year and outcome

decade_data <- termination_data %>% 
  filter(c_outcome_simple != 'ongoing') %>%
  filter(type_of_conflict == 2 | type_of_conflict == 4) %>%
  mutate(decade = (year %/% 10) * 10) 

# Create the line chart
ggplot(decade_data, aes(x = decade, fill = c_outcome_simple)) +
  geom_bar(position = "stack") +
  scale_x_continuous(breaks = seq(from = 1900, to = 2020, by = 10)) +
  labs(
    title = "War Terminations Over Time by Outcome",
    x = "Year",
    y = "Number of Wars Terminated",
    color = "Outcome Type"
  ) +
  theme_minimal()
```
# Outcomes of Interstate Wars

```{r}
# Pre-calculate the counts per year and outcome
decade_data <- termination_data %>% 
  filter(c_outcome_simple != 'ongoing') %>%
  filter(type_of_conflict == 2 ) %>%
  mutate(decade = (year %/% 10) * 10)
```


```{r}
# Get counts
decade_data
```


```{r}
# Create the line chart
ggplot(decade_data, aes(x = decade, fill = c_outcome_simple)) +
  geom_bar(position = "stack") +
  scale_x_continuous(breaks = seq(from = 1900, to = 2020, by = 10)) +
  labs(
    # title = "War Terminations Over Time by Outcome",
    x = "Decade",
    y = "Number of Wars Terminated",
    color = "Outcome Type",
    fill = "Means of War End"
  ) +
  theme_minimal() +
  theme(
    # Move legend inside the plot area (top right)
    legend.position = c(0.98, 0.98),
    # Anchor the legend by its top-right corner so it doesn't bleed off the edge
    legend.justification = c("right", "top"),
    # Optional: Add a background/border to ensure readability against the bars
    legend.background = element_rect(fill = "white", color = "black", linewidth = 0.2),
    axis.title = element_text(size = 11, color = "black"),
    axis.text = element_text(size = 11, color = "black")
  )
```
```{r}
# total number of conflicts
termination_data %>% filter(type_of_conflict == 2) %>% select(conflict_id) %>% unique() %>% nrow()

```

```{r}
decade_count <- termination_data %>% 
  filter(c_outcome_simple != 'ongoing') %>%
  filter(type_of_conflict == 2) %>%
  count(decade, c_outcome_simple)
```

```{r}
# count of interstate conflict endins for whole dataset

decade_count %>%
  group_by(c_outcome_simple) %>%         
  summarize(total_n = sum(n))
```


```{r}

termination_data %>% 
  filter(c_outcome_simple != 'ongoing') %>%
  filter(type_of_conflict == 2) %>%
  group_by(c_outcome_simple, conflict_resumes) %>%
  summarize(total_n = n(), .groups = 'drop')

```


```{r}
termination_data %>% 
  filter(c_epterm == 1) %>%
  filter(c_outcome_simple != 'ongoing') %>%
  filter(type_of_conflict == 2) %>%
  filter(conflict_resumes == 1) %>%
  group_by(c_outcome_simple) %>%         
  summarize(
    count = n(),
    avg_peace = mean(peace_duration, na.rm = TRUE),
    .groups = "drop" # Keeps the output as a clean, ungrouped tibble
  )
```
```{r}
termination_data %>% 
  filter(c_epterm == 1) %>%
  filter(c_outcome_simple != 'ongoing') %>%
  filter(type_of_conflict == 2) %>%
  filter(conflict_resumes == 1) %>% 
  select(c_outcome_simple, peace_duration) %>% 
  arrange(peace_duration)
  )
```
```{r}
termination_data %>% 
  filter(c_epterm == 1) %>%
  filter(c_outcome_simple != 'ongoing') %>%
  filter(type_of_conflict == 2) %>%
  filter(conflict_resumes == 1) %>%
  group_by(c_outcome_simple, peace_fails_in_10) %>%         
  summarize(
    count = n(),
    avg_peace = mean(peace_duration, na.rm = TRUE),
    .groups = "drop" # Keeps the output as a clean, ungrouped tibble
  )
```
```{r}
termination_data %>% 
  filter(c_epterm == 1) %>%
  filter(c_outcome_simple != 'ongoing') %>%
  filter(type_of_conflict == 2) %>%
  group_by(c_outcome_simple, peace_fails_in_10) %>%         
  summarize(
    count = n(),
    avg_peace = mean(peace_duration, na.rm = TRUE),
    .groups = "drop" # Keeps the output as a clean, ungrouped tibble
  )
```

```{r}
termination_data %>% 
  filter(c_epterm == 1) %>%
  filter(c_outcome_simple != 'ongoing') %>%
  filter(type_of_conflict == 2) %>%
  group_by(c_outcome_simple) %>%         
  summarize(
    count = n(),
    percent_fail = mean(peace_fails_in_10),
    .groups = "drop" # Keeps the output as a clean, ungrouped tibble
  )
```





# Outcomes of Interstate and Internationalized Wars

```{r}
# Create the line chart
ggplot(termination_data %>%
    filter(c_outcome_simple != 'ongoing') %>%
    filter(type_of_conflict %in% c(2,4)), 
  aes(x = decade, fill = c_outcome_simple)) +
  geom_bar(position = "stack") +
  scale_x_continuous(breaks = seq(from = 1900, to = 2020, by = 10)) +
  labs(
    title = "War Terminations Over Time by Outcome",
    x = "Year",
    y = "Number of Wars Terminated",
    color = "Outcome Type"
  ) +
  theme_minimal()
```

```{r}
decade_plus_count <- termination_data %>% 
  filter(c_outcome_simple != 'ongoing') %>%
  filter(type_of_conflict %in% c(2,4)) %>%
  count(decade, c_outcome_simple)

decade_plus_count %>%
  filter(decade > 1999) %>%     
  group_by(c_outcome_simple) %>%         
  summarize(total_n = sum(n))
```

```{r}
decade_plus_count %>%
  group_by(c_outcome_simple) %>%         
  summarize(total_n = sum(n))
```

# Duration of peace


```{r}
restart_analysis <- termination_data %>%
  filter(type_of_conflict %in% c(2,4)) %>%
  filter(c_outcome_simple %in% c("diplomatic", "military")) %>%
  filter(c_epterm == 1) %>%
  
  # Group by the type of termination
  group_by(c_outcome_simple) %>%
  
  # Calculate the metrics
  summarize(
    total_terminations = n(),
    num_restarts = sum(!is.na(duration)),
    avg_peace_duration = mean(duration, na.rm = TRUE)
  )

print(restart_analysis)
```

```{r}
restart_analysis <- termination_data %>%
  filter(type_of_conflict %in% c(2,4)) %>%
  filter(decade > 1999) %>%     
  filter(c_outcome_simple %in% c("diplomatic", "military")) %>%
  filter(c_epterm == 1) %>%
  
  # Group by the type of termination
  group_by(c_outcome_simple) %>%
  
  # Calculate the metrics
  summarize(
    total_terminations = n(),
    num_restarts = sum(!is.na(duration)),
    avg_peace_duration = mean(duration, na.rm = TRUE)
  )

print(restart_analysis)
```




# Scrap 

```{r}
ggplot(termination_data %>% 
         filter(c_outcome_simple != 'ongoing') %>% 
           filter(type_of_conflict == 4), 
       aes(x = year, fill = c_outcome_simple)) +
  geom_bar(position = "stack") +
  scale_x_continuous(
    breaks = seq(from = 1900, to = 2020, by = 10) # Adjust start/end to your data
  ) +
  labs(
    title = "War Terminations Each Year by Outcome",
    x = "Year",
    y = "Wars Terminated",
    fill = "Outcome Type"
  ) +
  theme_minimal()

```


```{r}
ggplot(termination_data %>% 
         filter(c_outcome_simple != 'ongoing') %>% 
           filter(type_of_conflict == 2), 
       aes(x = year, fill = c_outcome_simple)) +
  geom_bar(position = "stack") +
  scale_x_continuous(
    breaks = seq(from = 1900, to = 2020, by = 10) # Adjust start/end to your data
  ) +
  labs(
    title = "War Terminations Each Year by Outcome",
    x = "Year",
    y = "Wars Terminated",
    fill = "Outcome Type"
  ) +
  theme_minimal()
```



```{r}
# Pre-calculate the counts per year and outcome

plot_data <- termination_data %>% 
  filter(c_outcome_simple != 'ongoing') %>%
  count(year, c_outcome_simple)

# Create the line chart 
ggplot(plot_data, aes(x = year, y = n, color = c_outcome_simple)) +
  geom_line(linewidth = 1) +
  scale_x_continuous(
    breaks = seq(from = 1900, to = 2020, by = 10)
  ) +
  labs(
    title = "War Terminations Over Time by Outcome",
    x = "Year",
    y = "Number of Wars Terminated",
    color = "Outcome Type"
  ) +
  theme_minimal()
```


```{r}
# Pre-calculate the counts per year and outcome
plot_data <- termination_data %>% 
  filter(c_outcome_simple != 'ongoing') %>%
  filter(type_of_conflict == 2) %>%
  count(year, c_outcome_simple)

# Create the line chart

ggplot(plot_data, aes(x = year, y = n, color = c_outcome_simple)) +
  geom_line(linewidth = 1) +
  scale_x_continuous(
    breaks = seq(from = 1900, to = 2020, by = 10)
  ) +
  labs(
    title = "War Terminations Over Time by Outcome",
    x = "Year",
    y = "Number of Wars Terminated",
    color = "Outcome Type"
  ) +
  theme_minimal()
  
```


